\documentclass[a4paper,twoside,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{indentfirst}
\usepackage{fancyvrb}
\usepackage[left=2cm,top=2cm,right=2cm,nohead]{geometry}
\usepackage{hyperref}
\pagestyle{plain}

\makeatletter
\renewcommand\@seccntformat[1]{\csname the#1\endcsname.\quad}
\renewcommand\numberline[1]{#1.\hskip0.7em}
\makeatother

\author{Krzysztof Kwaśniewski}
\title{\LARGE Documentation for PyAgent}

\begin{document}
	\maketitle

	\tableofcontents
	\newpage

	\section{Summary}
		PyAgent is an agent for RHQ, an enterprise management system. It provides RHQ\footnote{\href{http://rhq-project.org/display/RHQ/Home}{http://rhq-project.org/display/RHQ/Home}} with data gathered from desktop stations scattered across the network so that the stations could be monitored and managed. The data RHQ can handle can be of various types, one refers to them as \emph{metrics}. These may include for instance repetitive measurements of a CPU's temperature over the time.

		To collect this data, PyAgent communicates with CIM servers, which are WBEM\footnote{\href{http://dmtf.org/standards/wbem}{http://dmtf.org/standards/wbem}}-based services. The WBEM standard describes an abstract interface that allows to interact with different computer systems and - for instance - access a data on a aforementioned CPU's temperature (and many more).

		PyAgent is extensible and it's possible to quite easily add to it support for more protocols - so that it could gather the data from different sources, not only WBEM-based, and send it to a different destination, not only RHQ. Although now there does not seem to be an urgent need for that, one day it may turn out to be necessary.

		This project has been started in April/May 2012 as part of The Google Summer Of Code program, mentored by Heiko Rupp\footnote{\href{https://community.jboss.org/people/pilhuhn}{https://community.jboss.org/people/pilhuhn}}.
		
		The code is implemented in Python 2.x; will not work with Python 3.x, since it depends on a python-wbem library to communicate with WBEM-based services. The following list is - I hope - a full list of PyAgent's dependencies:

		\begin{itemize}
			\item python-wbem\footnote{\href{http://pywbem.sourceforge.net/}{http://pywbem.sourceforge.net/}}
			\item python-nose\footnote{\href{http://nose.readthedocs.org/en/latest/}{http://nose.readthedocs.org/en/latest/}}
			\item python-coverage\footnote{\href{http://nedbatchelder.com/code/coverage/}{http://nedbatchelder.com/code/coverage/}}
			\item python-lxml\footnote{\href{http://lxml.de/}{http://lxml.de/}}
			\item python-distutils\footnote{\href{http://pypi.python.org/pypi/setuptools/0.6c11}{http://pypi.python.org/pypi/setuptools/0.6c11}}
			\item make
			\item rpmbuild (to build an RPM package for Fedora)
			\item pdflatex (to produce this documentation)
		\end{itemize}

		This is a free software licensed under the terms of GNU General Public License version 3 or later as published by the Free Software Foundation.

	\section{Running the code}
		To run the code, you need to invoke the \texttt{rhqagent} startup script located in the top project's directory. As a minimum, you need to provide an argument \texttt{-{}-configuration-file} pointing at an xml configuration file, as PyAgent needs to know where it can gather the data from and where should it send it to.

		PyAgent can help you generate such a configuration file, if you also pass a \texttt{-{}-setup} argument to it.

		For more details, please invoke PyAgent with a \texttt{-{}-help} argument.

		You can also build packages, currently an RPM and in the Python Package Index format, so that the code could be distributed easier. After installing such a package, you will be also able to invoke PyAgent with an init script \texttt{rhqagentd}. To do that (at least in both Ubuntu and Fedora), invoke a command \texttt{service rhqpyagentd} to get a list of supported arguments.

	\section{Directory structure}
		To give you an overall impression on how this project is organised, here you have a brief description of the files it consists of.

		\begin{itemize}
			\item COPYING -- contains the text of GPLv3 license under which (or any later) this project is licensed.
			\item docs -- the documentation of PyAgent lives here in a TEX format. To turn it into PDF, run \emph{make docs} in the main project's directory.
			\item Makefile -- lets one perform a couple of actions. One can use it to run unit tests with or without code coverage measurement; create a PDF file with documentation out of this one in a TEX format; build an RPM or a PyPi package and - finally - clean up all the temporary files created during the previous operations.
			\item packages -- RPM and PyPi packages are created in this directory.
			\item pyagent -- contains the source code of the application.
				\begin{itemize}
					\item agent -- a main module of PyAgent that governs all the operations it performs.
					\item data
						\begin{itemize}
							\item model -- PyAgent's data model.
							\item provider -- the abstract definition for classes that are able to provide data for PyAgent and all the implementations derived from it.
							\item sender -- the abstract definition for classes that are able to pass further data gathered by PyAgent and all the implementations derived from it.
						\end{itemize}
					\item exception -- contains definitions of exceptions thrown by PyAgent's submodules.
					\item \_\_init.py\_\_ -- necessary to turn this directory into a Python module (available for importing).
					\item settings -- a module responsible for generating and parsing the configuration file as well as parsing the arguments passed in the command line.
						\begin{itemize}
							\item sample\_settings.xml -- sample configuration file. If you want to define your own configuration file by hand, here you will find an example of such a file.
							\item settings.xsd -- an xml schema for PyAgent's configuration file. The configuration file must comply with this schema.
						\end{itemize}
					\item test -- unit tests live here.
				\end{itemize}
			\item README -- a brief description of the project.
			\item rhqpyagent -- a startup script for PyAgent.
			\item rhqpyagentd -- an init script that turns PyAgent to behave as a daemon (system service), this is the expected way of running PyAgent after its RPM or PyPi package has been installed.
			\item setup.py -- necessary to build a PyPi package, you don't need to use it directly, use Makefile instead.
		\end{itemize}

	\section{How to...?}
		The syntax of the configuration file is somewhat self-explanatory and in the pyagent.settings submodule there is a sample configuration file. I won't go into deep details here, as it's not very complicated. On the other hand a couple of things need to be written down.

		\subsection{Add a new data provider/sender}

			As mentioned before, PyAgent was designed in an extensible way. You can quite easily add to it a new data provider (let's focus on that, data senders work in the same way). You do this by subclassing pyagent.data.provider.AbstractDataProvider.AbstractDataProvider. No, this is not a  mistake. The first AbstractDataProvider refers to the module name and the class named in the same way is stored inside.

			Your implementation has to be called according to the schema \texttt{<<PROTOCOL>>DataProvider} and live in a file \texttt{pyagent.data.provider.implementation.<<PROTOCOL>>DataProvider}, where \texttt{<<PROTOCOL>>} is an uppercase name of the protocol for which you're just adding a new provider. For instance, to support FTP, you would create a class FTPDataProvider in pyagent.\-data.\-provider.\-implementation.\-FTPDataProvider. After that, you can put \emph{ftp} as a protocol name in the configuration file and PyAgent will understand what your intentions are.

		\subsection{Add a callback}

			What is a callback and why it there a need for it? At times you might want not only to pass raw measurements' data to RHQ, but also to first prepare it in some way. Let's consider a couple of examples -- well, maybe some of them are not very realistic, but just to give you an overall impression:
			\begin{itemize}
				\item Instead of analyzing the amount of free space on your hard disk, you might want to compute a derivative of it over the time and analyze this \emph{artificial metric}.
				\item You might want to measure some kind of an unreliable metric, burden with huge errors. Let's assume its value in fact cannot ever go over 100, however when the real value and an error add up, then it's often higher and you prefer to rather return 100 in such a case. So you slightly modify your original metric and you are rather monitoring an \emph{artificial}, another one.
				\item The system you are monitoring might return values expressed in a 7-based numbering system and you would need to convert them to a 11-based one for some reason. So, just like in the previous examples, you produce an \emph{artificial metric} for that.
			\end{itemize}
			Obviously, there are many more scenarios. The point is: you need to define a convert function that will compute values of the artificial metric that fulfills your requirements.
				
			So, basically, a callback is a Python function that you define and that will be executed by PyAgent over and over to compute the value for an artificial metric. The configuration file generator included in PyAgent does not help creating callbacks, as they are quite an advanced issue. You need to define them on your own, by hand.

			A callback definition consists of three things. The first one is wrapped up with \texttt{function} tags and it defines the Python function mentioned above, called from time to time to compute the metric's value. The second one is a specification of parameters passed to this function and the last part is a list of destination servers' mappings -- once you have your metric computed, what should PyAgent do with it? You need to give it guidance on where to send the measurements.

			Let's see an example:
			\begin{Verbatim}
<callbacks>
	<callback>
		<function>
		<![CDATA[
metric1 = params[0]
metric2 = params[1]

return metric1[0].value+0.5*metric1[1].value \
	+ metric2[0].value+0.5*metric2[1].value
		]]>
		</function>
			\end{Verbatim}

   			A callback, as already mentioned, is a Python function:
			\begin{Verbatim}
def fun(params):
	[[CODE]]
			\end{Verbatim}
			and the user is the one responsible for providing the contents for [[CODE]]. \textbf{The code for callback you'll write must be aligned to the beginning of the line - pay attention to the line with an assignment: metric1 = params[0]. The callback wouldn't work with any space prepended to this statement. It must start at the beginning of the line.}

			When writing your own callback you are provided with a variable \texttt{params} which is a list of arguments that you specify in the configuration file after the callback's source code. Consider the example shown above where two arguments are pulled from params. Now what are actually \texttt{metric1} and \texttt{metric2}? Well, those are also lists - lists of measurements of appropriate metrics taken in the past. By default 100 of them are available, the capacity of past values registry can be changed at command line, when starting the agent up. The newest measurement is kept at index 0 with increasing indices supplying older values.

			For each measurement taken, that is \texttt{params[i][j]}, you can use two properties. The first, \texttt{value}, is the actual value of the measurement. The second, \texttt{timestamp}, is a floating point number indicating the time of having taken the measurement.

			Summing up the things written above, you would write \texttt{params[3][1].timestamp} to access the time when the penultimate measurement of the fourth argument of the callback was taken. You should not rely on any behaviour of callbacks that is not defined here.

\end{document}

